---
title: "WUP_04"
author: "Jake Glickman"
date: "June 2, 2022"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
GSScsv <- read.csv("GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(GSScsv,aes(x=DEGREE,y=WORDSUM))

basicCN + geom_boxplot()
```

The graph shows that there is a difference in Wordsum test scores between grade levels. The LT High School box is below the Junior College and the Graduate degrees.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSScsv)
```

The mean of graduate students is 2.25 above the mean of LT High School students. The Q1 of graduate students is 10, which is larger than the Q3 of High School students which is 9.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSScsv)
summary.aov(model1)
```

The chances of getting results such as this is there was no difference between subgroups is 2.36 out of 10,000. This leads us to believe there is a difference in the test scores between subgroups.

Now that we have a low P value we want to see what values are different from one another. We will now use a tukey test to find the differences.

```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 16.7, 2, 2))
plot(TukeyHSD(model1),las=2)
```

Any confidence intervals that do not reach the 0 are different. LT High School compared to Graduate is shown to be different because it does not reach the 0. High School compared to Graduate is different because it does not reach the 0 either.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The mean of LT High School students compared to the mean of Graduate students is 2.25 lower. The first inferential test showed us that the chances of getting these results if there was no difference is a 2.36 out of 10,000 chance. Finally we compared the results of grade levels through a tukey test which showed there was a difference between High School and Graduate results and LT High School and Graduate results. Neither of the confidence intervals reached 0.
